---
title: "Zoning with GeoFIS"
subtitle: "Package GeoFIS `r packageVersion('GeoFIS')`"
author: "Jean-Luc Lablée, Serge Guillaume"
output: 
  rmarkdown::html_vignette:
    fig_height: 5.5
    fig_width: 7
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Zoning with GeoFIS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: "`r system.file('REFERENCES.bib', package='GeoFIS')`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette shows how to generate management zones from data with GeoFIS [@Pedroso10].

See [Zoning documentation](https://www.geofis.org/en/documentation-en/zoning/) for more details about the zoning algorithm.

```{r, setup}
library(GeoFIS)
```

## Loading data

The zoning process accepts input dataset of class [SpatialPointsDataFrame](https://www.rdocumentation.org/packages/sp/topics/SpatialPointsDataFrame) or [SpatialMultiPointsDataFrame](https://www.rdocumentation.org/packages/sp/topics/SpatialMultiPointsDataFrame) of the [sp](https://cran.r-project.org/package=sp) R package.<br>
It keeps only the attributes that can be used in the zoning process, meaning numeric attributes, without missing values and with a range that is not limited to a unique value. The last condition is required by the min-max standardization process.<br>
The input dataset must be in a projected coordinate reference system.

In this example the `conductivity_2014` input dataset is used. It is included in the GeoFIS R package, each location is described by a single numeric attribute `conduct`:

```{r}
data(conductivity_2014)
zoning <- NewZoning(conductivity_2014)
```

The zonable dataset used by the zoning process is available through the `zonable_data` method. In this example `zonable_data` is identical to `conductivity_2014`.

```{r}
zonable_data <- zoning$zonable_data()
```

## Main parameters

### The border

The border is a polygon used to limit the processed area. Only data points within the border are processed.

If `NULL` the Convex Hull polygon of the input dataset will be used (computed with [gConvexHull](https://www.rdocumentation.org/packages/rgeos/topics/gConvexHull) function of the [rgeos](https://cran.r-project.org/package=rgeos) R package).

```{r}
zoning$border <- NULL
```

An object of class [SpatialPolygons](https://www.rdocumentation.org/packages/sp/topics/SpatialPolygons) can be used as border:

```{r}
data(conductivity_border)
zoning$border <- conductivity_border
```

#### Show the Voronoi diagram

```{r}
zoning$perform_voronoi()
vm <- zoning$voronoi_map()
par(mar = c(0, 0, 0, 0))
plot(vm)
plot(zoning$zonable_data(), add = TRUE)
```

### The neighborhood

The neighborhood relation can be filtered by a minimal common edge length.

If `NULL` all contiguous Voronoi polygons are considered as neighbors: 

```{r}
zoning$neighborhood <- NULL
```

The user can define the minimum edge length (in the same unit as the input dataset coordinate system) shared by two Voronoi polygons for being considered as neighbors:

```{r}
zoning$neighborhood <- 5
```

#### Show the neighborhood relations

The red lines shows the pairs of Voronoi polygons that are no longer considered as neighbors.

```{r}
zoning$perform_neighborhood()
nm <- zoning$neighborhood_map()
par(mar = c(0, 0, 0, 0))
plot(vm)
plot(subset(nm, filtered == FALSE), add = TRUE, col = "green")
plot(subset(nm, filtered == TRUE), add = TRUE, col = "red")
```

### Attribute distance

This distance function is used for a given attribute. It computes the distance between two data points in the monodimensional attribute space. As many distance functions as there are attributes in the zonable dataset are needed (in a list if multiple attributes).<br>
Two univariate distances are available: `EuclideanDistance` (the default) or `FuzzyDistance`. Or `NULL` if the attribute should not be used in the zoning process.

The fuzzy distance function is based on a fuzzy partition that allows for integrating expert knowledge into distance calculations [@Infsci13; @fuzzieee13]. The partition must be a standardized fuzzy partition based on a `FisIn` object of the [FisPro](https://cran.r-project.org/package=FisPro) R package.

The following command-line sets the fuzzy distance based on 3 Mfs-partition defined by the following breakpoints: 20, 30, 100. The range of the `conduct` attribute is [12, 116]:

```{r}
zoning$attribute_distance <- FuzzyDistance(NewFisIn(c(20, 30, 100), 12, 116))
```

The default value, used in this example, is the euclidean distance:

```{r}
zoning$attribute_distance <- EuclideanDistance()
```

### Zone distance aggregation

To compute the distance between two zones, all the data points included in the two zones are considered and the aggregation is done using the aggreg parameter: 

$$ d(z_i,z_j) = (Aggreg) d(x,y), \forall x \in z_i, y \in z_j$$
Three aggreg operators are available: `MinimumDistance`, `MaximumDistance` (the default) or `MeanDistance`.<br>
The two zones to be merged at a given iteration are the ones for which the zone distance is minimum.

```{r}
zoning$zone_distance <- MaximumDistance()
```

### Combine distance

The combination function distance is needed when the zoning is done according to several attributes. In this case, each univariate or elementary distance is computed and normalized in a unit interval. These partial distances are then aggregated to yield the distance between two data points in the multidimensional attribute space. The distance combination is done before computing the between-zones distance.<br>
Two combinations are proposed: `EuclideanDistance` (the default) or `MinkowskiDistance`.

```{r}
zoning$combine_distance <- EuclideanDistance()
```

### Smallest zone

This criterion is used to determine the smallest size for a zone (number of points or area) to be kept in the final map. The zones with a size less than the threshold are included in the most compatible neighboring zone.<br>
The two available parameters are: `ZoneSize` or `ZoneArea`.<br>
The default value is `ZoneSize` with 1 point.

```{r}
zoning$smallest_zone <- ZoneSize(1)
```

## Perform zoning

```{r}
zoning$perform_zoning()
```

Get the map with 5 zones:

```{r}
map5 <- zoning$map(5)
```

The result map contains for each zone:

* the id
* the number of points
* the area
* the mean of the attribute
* the std of the attribute

```{r}
print(map5@data)
```

Plot the map:

```{r}
library(RColorBrewer)

palette <- brewer.pal(3, "Blues")
breaks <- c(30, 50, 70, 100)
colors <- palette[findInterval(map5$conduct_mean, vec = breaks)]
par(mar = c(0, 0, 2, 0))
plot(map5, col = colors, main = "Map 5 zones")
legend("bottomright", legend = levels(cut(map5$conduct_mean, breaks)), 
  fill = palette, title = "Conductivity mean")
```

### Save the map

Any map can be exported to various common file format, e.g. Shapefile:

```{r, message=FALSE, warning=FALSE}
library(rgdal)

writeOGR(map5, dsn = ".", layer = "map5", driver = "ESRI Shapefile")
```

```{r, include=FALSE}
file.remove(dir(path = ".", pattern = "map5.*"))
```

## Publications

---
nocite: |
  @agriculture8060073
...
